---
title: "Merge and clean penguins data"
author: "Andrea"
date: "2025-01-23"
knit: resrepo::knit_to_results
output: md_document
---

```{r setup_rmd, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

It is useful to set your input directories all at the top of your script, so that
you can easily check in one place what is needed to run it. 
You should always use relative paths, either relative
to the script or, if you use `resrepo::path_resrepo`, to the project root.
```{r input_paths}
tux_dir <- "../data/raw/original"
# alternatively:
# raw_dir <- path_resrepo("/data/raw/original")
palmer_dir <- "../data/raw/s01_download_penguins"
```

And the output paths:
```{r output_paths}
output_dir <- "../intermediate/s02_merge_clean"
dir.create(output_dir, showWarnings = FALSE)
```

Note that, by using "showWarnings = FALSE", we suppress the warning that the directory already exists. This is useful when the script is run multiple times, as it will not stop the script.

## Reading data

We can now read the data, using the paths that we set up at the beginning of our script:
```{r read_data}
tux_penguins <- read.csv(file.path(tux_dir,"tux_measurements.csv"))
palmer_penguins <- read.csv(file.path(palmer_dir, "palmer_penguins.csv"))
```

## Merge and clean

We will now merge the two datasets, and remove lines with missing data:
```{r merge_clean}
# merge the two datasets
penguins <- rbind(tux_penguins, palmer_penguins)
# remove na
penguins_na_omit <- na.omit(penguins)
```

## Saving the data

We will now save the data into the appropriate `/data/intermediate` directory:
```{r save_data}

write.csv(penguins_na_omit, file.path(output_dir, "penguins_na_omit.csv"))
```
