---
title: "Data versioning"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Data versioning}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
example_dir <- file.path(tempdir(),"resrepo_example")
unlink(example_dir,recursive = TRUE)
dir.create(example_dir, showWarnings = FALSE)
example_repo <- git2r::init(example_dir)
git2r::config(example_repo, user.name = "Test", user.email = "test@example.org")
vignette_dir<-getwd()
knitr::opts_knit$set(root.dir = example_dir)
# setwd(example_dir)
```

In many scientific projects, data are simply too large to be stored in a git repository. 
However, you still want to keep track of the data you used for my analysis, and often you will
be using different code branches to explore alternative approaches (e.g. different data filtering).
`res_repo` allows you to create versions of the data directories used in your projects, 
allowing you to easily switch between different versions of the data.

# Setting up data versioning

Let us start by recreating the example repository. This is the same repository
we created in vignette basic_workflow; if you still have that repository, you don't need to run the code below:

```{r}
library(resrepo)
init_resrepo()
file.copy(from = system.file("vignette_example/tux_measurements.csv", package="resrepo"),
          to = path_resrepo("/data/raw/original/tux_measurements.csv"),
          overwrite = TRUE)
file.copy(from = system.file("vignette_example/s01_download_penguins.Rmd", package="resrepo"),
          to = path_resrepo("/code/s01_download_penguins.Rmd"),
          overwrite = TRUE)
knit_to_results(path_resrepo("/code/s01_download_penguins.Rmd"))
file.copy(from = system.file("vignette_example/s02_merge_clean.Rmd", package="resrepo"),
          to = path_resrepo("/code/s02_merge_clean.Rmd"),
          overwrite = TRUE)
knit_to_results(path_resrepo("/code/s02_merge_clean.Rmd"))
file.copy(from = system.file("vignette_example/s03_pca.Rmd", package="resrepo"),
          to = path_resrepo("/code/s03_pca.Rmd"),
          overwrite = TRUE)
knit_to_results(path_resrepo("/code/s03_pca.Rmd"))
git2r::add(path=".")
git2r::commit(message="Save plot", all=TRUE)

```
Let us check that we do have indeed a full repository:
```{r echo=FALSE}
fs::dir_tree()
```

If our data become too large to track on git, and we want to start versioning the data, we can use the function `setup_version`.
```{r}
version_setup(quiet = TRUE)

```

Let us check what happened:
```{r echo=FALSE}
fs::dir_tree()
```

We can see that both `data/raw` and `data/intermediate` have been moved into `version_resources`, under the `initial` version,
and that they have been replaced by symlinks. If we use these links as if they were directories, they point correctly
to the directories in `version_resources`.

```{r}
dir(path_resrepo("/data/raw"))
```

However, note that the data are now NOT being tracked by git, as `version_resources` is in .gitignore. So,
you want to make sure that you have a reliable way to back up your data (we provide a solution
in the vignette on `Data mirroring outside a git repository`).

# Adding a new version
We might now want to revise the filtering of data that we are doing. As this will
affect some intermediate and results, we want to version the data before we start,
so that we can compare the two versions.
```{r}
version_add(new_version = "new_filtering",
            description = "Filtering out some data")
```
```{r echo=FALSE}
fs::dir_tree()
```

If we write a file to the `data/raw` directory, it will be written to the `version_resources/new_filtering/raw` directory.
```{r}
write.csv("blah", path_resrepo("/data/raw/original/my_new_file1.csv"))
```

```{r echo=FALSE}
fs::dir_tree()
```

# Switching between versions
We can now switch between the versions by using the `version_switch` function. TO BE WRITTEN!!!!
```{r}
version_switch("initial")
```

Check that we are indeed writing to the `initial` version:
```{r}
write.csv("blah", path_resrepo("/data/raw/original/my_new_file2.csv"))
```

```{r echo=FALSE}
fs::dir_tree()
```
