---
title: "Sharing a resrepo repository"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Sharing a resrepo repository}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
---

This vignette covers how to share a resrepo repository. This could be sharing
with collaborators (or anyone planning to work on it), or sharing it with the
public (e.g. as part of a publication). 

If sharing with collaborators, to use it, they will simply need to:

1. Clone the resrepo repository
2. Obtain the data they need
3. And set up symlinks to the location of the data

If sharing with the public or with reviewers, in addition, you will need to 
upload the data to a publicly available location (e.g. Zenodo, Dryad, OSF, 
etc..) and provide a script to download the data.

## 1 & 2. Cloning a resrepo repository and downloading data

@TODO: Clone the resrepo_penguins repository - create gif

```{r setup, echo=FALSE, results='hide'}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)

library(resrepo)

directory_penguins <- file.path(tempdir(), "resrepo_penguins")
if (!dir.exists(directory_penguins)) dir.create(directory_penguins)

# clone the penguin repository (public)
git2r::clone("https://github.com/EvolEcolGroup/resrepo_penguins",
             local_path = directory_penguins,
             branch = "main",
             progress = TRUE)

knitr::opts_knit$set(root.dir = directory_penguins)

```


So you have cloned a resrepo repository with code in it, but the data to run the 
analyses are stored somewhere else outside the cloned repository.
```{r}
fs::dir_tree()
```


As you can see from the above tree, `data` is an empty folder,  this is because 
no data are stored in the repository. 

Therefore, you need to take two steps:

- Download the data from where they are stored and place them in a location of your choice (e.g. a folder on 
   your local machine, a synced folder, etc..)
- Set up the symlinks to point to the location of the data you just downloaded

Moreover, as you can see, this repository contains the folder `version_meta` which indicates to us that this
repository has been versioned. If you are not familiar with the concept of versioning,
please check the  `data versioning` vignette. 

As an example, we will create a directory to act as our external data storage folder. Note that we we are creating this folder in the temporary directory, but you should choose a location anywhere on your local machine. We will call this folder `external_data_storage_new`.
```{r}
# create new external folder for data 
external_data_storage_new <- file.path(tempdir(), "external_data_storage_new")
#unlink(external_data_storage_new, recursive = TRUE)
dir.create(external_data_storage_new)
```

It is important that the external data folder where you store your data is structured in the exact same way as your `data` folder within your resrepo repository. Otherwise `resrepo` will not be able to find the files.

Ideally, the directory structure of the data shared with you for this project should already match the structure of the `data` folder in the repository. Check with the person who shared the repository with you if you are unsure about this.

In the following example, we will create the substructure of the `data` folder, which contains `starting/raw` and `initial/intermediate` directories, 

```{r}
intermediate_path <- file.path(tempdir(), "external_data_storage_new/initial/intermediate")
raw_path <- file.path(tempdir(), "external_data_storage_new/starting/raw")
#unlink(intermediate_path, recursive = TRUE)
#unlink(raw_path, recursive = TRUE)
dir.create(intermediate_path, recursive = TRUE)
dir.create(raw_path, recursive = TRUE)
```

And then we will copy the raw data into `external_data_storage_new/starting/raw`:
```{r}
file.copy(from = system.file("vignette_example/tux_measurements.csv", package="resrepo"),
          to = raw_path,
          overwrite = TRUE)
```

## 3. Setting up symlinks to data

If you have a repository that is not versioned (it does not contain any version.meta), then you can simply link your external data storage location to the `data` folder in your repository using the `data_dir_link` function:

Because this repository has been versioned, we need to set up versioning in our 
local repository. Now that our `external_data_storage_new` folder contains the data, we can do this
using `version_setup`. 

We need to supply the argument `resources_path` to `version_setup`, which is the path to the external data storage folder:
```{r}
version_setup(quiet = TRUE, resources_path = external_data_storage_new)
```


```{r}
fs::dir_tree()
```

We can now see that we have `data/intermediate` and `data/raw` folders. These are both links which point to the `versions` folder, and `versions` points to our external data storage (easy isn't it!);

  data/intermediate -> versions -> external_data_storage_new/initial/intermediate
  data/raw -> versions -> external_data_storage_new/starting/raw

We can also check the structure of the `external_data_storage_new` folder:
```{r}
fs::dir_tree(external_data_storage_new)
```

## 4. Downloading data from a public location

If you are sharing your repository with the public, you will need to upload your 
data to a publicly available location (e.g. Zenodo, Dryad, OSF, etc..) and provide 
a script to download the data. You may also require data from other public repositories (e.g. GBIF, GenBank, etc..).
The script to download the data should be placed in the `code` folder of your resrepo repository, and should
ideally be named `s0_download_data.Rmd`. It should download all the data from the
different sources and place them in the appropriate directory (see Section 2 above).

After doing so, you can then use `version_setup` to set up the symlinks to the data as shown in Section 3 above.

You can find an example of an `s0_download_data.Rmd` in the `inst/vignette_example` folder of the `resrepo` package.